---
title: "CStart sample"
output: html_notebook
---

```{r}
library(tidyverse)
library(patchwork)
```

```{r}
source("CStart-functions.R")
```

Load the sample data set.
```{r}
data <- read_csv("CStart-sample.csv")
```
Plot the midline.
```{r}
data |> 
  ggplot(aes(x = mxmm, y = mymm, group = tsec)) +
  geom_path() +
  coord_fixed()
```
# Body width and height

We need the width and height of the body to estimate the center of mass, assuming that the body is constant density.

Read in the bluegill body shape file.
```{r}
shape <- read_csv("bluegillshape.csv") |> 
  mutate(s = pracma::linspace(0, 1, n()))
```
Show the body width and body height.
```{r}
p1 <- ggplot(shape, aes(x = s, y = widthL)) +
  geom_line()

p2 <- ggplot(shape, aes(x = s, y = heightL)) +
  geom_line()

p1 + p2 + plot_layout(ncol = 1)
```

# Estimate COM

## Stretched-straight COM

This first estimates the location of the stretched straight COM on the body,
based on the width and height, assuming constant density.
```{r}
sscomdist <- get_stretched_straight_com(shape$widthL, shape$heightL)
sscomdist
```

Now we interpolate that position on the midline throughout the C start.
```{r}
comdata3 <-
  data |> 
  group_by(tsec) |> 
  group_modify(\(df, key) interp_stretched_straight_com(df, mxmm,mymm, ssdist=sscomdist)) |> 
  mutate(method = 'ss')
```

## COM from volume

Get the time varying center of mass, based on the volume of the fish, again
assuming constant density.

```{r}
comdata1 <-
  data |> 
  group_by(tsec) |> 
  group_modify(\(df, key) get_com(df, mxmm, mymm, 
                                  width = shape$widthL, height = shape$heightL)) |> 
  mutate(method = 'volume')
```

## COM from horizontal projected area

Get the time varying center of mass, based on the horizontal projected area of the 
fish, assuming contsant height and constant density.

```{r}
comdata2 <-
  data |> 
  group_by(tsec) |> 
  group_modify(\(df, key) get_com(df, mxmm, mymm, 
                                  width = shape$widthL)) |> 
  mutate(method = 'area')
  
```

Merge the results and compare them
```{r}
comdata <- ungroup(comdata1) |> 
  bind_rows(ungroup(comdata2)) |> 
  bind_rows(ungroup(comdata3))

comdata |> 
  # filter(method == "volume") |> 
  ggplot(aes(x = comx, y = comy, color = method)) +
  geom_path()
```

# Smooth data and take derivatives


```{r}
datas <-
  data |> 
  dplyr::filter(pt == 2) |> 
  mutate(y1 = mypix,
         y1s = smooth_1_position(y1, samplerate, stage1dur = 0.04, endmode="none"))

p1 <- datas |> 
  ggplot(aes(x = tsec)) +
  geom_point(aes(y = y1), color = "black") +
  geom_path(aes(y = y1s), color = "blue") +
  xlim(-0.01, 0.08)
  
p2 <- datas |> 
  ggplot(aes(x = tsec)) +
  geom_point(aes(y = y1s-y1), color = "blue") +
  xlim(-0.01, 0.08) +
  ylab("Difference between\nsmoothed and original")

p1 + p2 + plot_layout(ncol = 1)
```

```{r}
data |> 
  dplyr::filter(pt %in% c(1, 5, 10, 15, 20)) |> 
  group_by(pt) |> 
  group_modify(~ smooth_positions(.x, mxpix, mypix, samplerate, stage1dur = 0.04)) |> 
  ggplot(aes(x = tsec)) +
  geom_point(aes(y = mypix, group = pt)) +
  geom_path(aes(y = mypix.s, color = pt)) +
  xlim(-0.01, 0.08)
```



```{r}
data |> 
  dplyr::filter(pt %in% c(1, 5, 10, 15, 20)) |> 
  group_by(pt) |> 
  group_modify(~ smooth_positions(.x, mxpix, mypix, samplerate, stage1dur = 0.04)) |> 
  mutate(vx = nderivative(tsec, mxpix),
         vxs = nderivative(tsec, mxpix.s)) |> 
  mutate(pt = factor(pt)) |> 
  ggplot(aes(x = tsec)) +
  geom_point(aes(y = vx, color = pt)) +
  geom_path(aes(y = vxs, color = pt)) +
  xlim(-0.01, 0.08)
```

```{r}
comdata |> 
  group_by(method) |> 
  mutate(comvelx = nderivative(tsec, comx),
         comvely = nderivative(tsec, comy),
         comvel = sqrt(comvelx^2 + comvely^2)) |> 
  ggplot(aes(x = tsec, y = comvel, color = method)) +
  geom_path() +
  xlim(-0.01, 0.08)
```

